# Data transformation

## Remove Columns

```{r message=FALSE, warning=FALSE, include=FALSE}

library(readr)
cash_information_raw <- read_csv("C:/Users/chaoy/Desktop/Fall2021/5702_EDAV/finalproject_data/cash_information.csv")

# drop columns
cash_information <- subset(cash_information_raw, 
                           select =-c(`Municipality`, `Police Report`, `DOT Reference Marker Location`, `Pedestrian Bicyclist Action`))
```


Here is the overview of columns information before cleaning. First, variable `Date` need to convert to class `Date` in format "Year-Month-Day". Second, variable `Event Descriptor` will be convert into two new variables to narrow down the information contains in single variable. Finally, all variables under type character will be transformed as factor. 

```{r echo=FALSE}
str(cash_information)
```

## Change Format of Variable `Date` 

Here is the view of variable `Date` before the transformation. 
```{r echo=FALSE}
print(cash_information$Date[1])
print(class(cash_information$Date))
```
Here is the view of variable `Date` after the transformation. 
```{r  echo=FALSE}
# transform variable `Date` to date format
cash_information$Date <- as.Date(cash_information$Date, format="%m/%d/%Y")
print(cash_information$Date[1])
print(class(cash_information$Date))
```

## Extract reason of crashing from variable `Event Descriptor`

### `Collision_Type`
Following it the current levels in `Event Descriptor`. For generalization, new variable `Collision_Type` is introduced to indicate the high-level collision type: "Collision With Fixed Object", "Collision With Unfixed Object" (will be shortened as "Collision With" in the table), and "Non-collision"

For generating new variable `Collision_Type`, two changes needed:
1. "Deer" can be included in "Animal, Collision With" 
2. "Other Pedestrian" can be included in "Other Pedestrian, Collision With"

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
levels(factor(cash_information$`Event Descriptor`))
```

```{r echo=FALSE}
cash_information <- cash_information %>% 
  mutate(`Event Descriptor` = ifelse(`Event Descriptor`=="Deer", "Animal, Collision With",`Event Descriptor`)) %>% 
  mutate(`Event Descriptor` = ifelse(`Event Descriptor`=="Other Pedestrian", "Other Pedestrian, Collision With",`Event Descriptor`))
```

After cleaning and extracting keywords from `Event Descriptor`, here are the levels for variable `Collision_Type`:
```{r echo=FALSE}

# reference code for analyzing only
event_descriptor <- levels(factor(cash_information $`Event Descriptor`))
collision_type <- c()
for (i in event_descriptor){
  collision_type <- c(collision_type, sub(".*,", "", i))
}

# actual code for generating new variable
cash_information <- cash_information %>% mutate(Collision_Type = sub(".*,","", `Event Descriptor`))
levels(factor(cash_information$Collision_Type))
```


### `Collision_Detail`
Besides the general collision type information, `Event Descriptor` also told us the collision details, which will be stored in new variable `Collision_Detail`. Following is the view of collision details before data cleaning. 
```{r echo=FALSE}
event_descriptor <- levels(factor(cash_information $`Event Descriptor`))
collision_type <- c()
for (i in event_descriptor){
  collision_type <- c(collision_type, sub(",.*", "", i))
}
levels(factor(collision_type))
```
In the above categories, "Other Pedestrian" and "Pedestrian" are represented the same collision type, so they can combine into the same class "Pedestrian". The following shows the final 30 levels in `Collision_Detail`.
```{r echo=FALSE}
cash_information <- cash_information %>% 
  mutate(Collision_Detail = sub(",.*","", `Event Descriptor`)) %>% 
  mutate(Collision_Detail = ifelse(Collision_Detail=="Other Pedestrian", "Pedestrian", Collision_Detail))

levels(factor(cash_information$Collision_Detail))
```


## Convert to Factor Level
Finally, all the variable type convert to factor level for our further data analysis. Here is the table sumamary:
```{r echo=FALSE}

cash_information[sapply(cash_information, is.character)] <- lapply(cash_information[sapply(cash_information, is.character)], as.factor)
str(cash_information)
```
