# Results

```{r message=FALSE, warning=FALSE, include=FALSE}
library(readr)
df <- read_csv("C:/Users/chaoy/Desktop/Fall2021/5702_EDAV/finalproject_data/cash_information_clean.csv")

#df <- read.csv("~/Downloads/cash_information_clean.csv")

#df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)], as.factor)
```

## When
```{r include=FALSE}
library(tidyverse)
library(dplyr)
library(lubridate)
library(vcd)
library(patchwork)
library(RColorBrewer)
```

### Time in Day
First, we explore when crashes happen the most, and the corresponding reasons.
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}
df_time <- df %>% select(Time, Year) %>%
  mutate(hour = hour(Time)) %>%
  group_by(hour, Year) %>%
  summarize(n = n()) %>%
  ungroup()

ggplot(df_time, aes(hour,n, fill=factor(Year))) + 
  geom_bar(stat="identity", position=position_dodge()) + 
  scale_fill_brewer(palette="Set2") +
  theme_minimal()
```
Here is an overview of number of crash occur in 24 hours in 2018 and 2019. Firstly, there are three peaks in the plot: 0:00, 8:00 and 17:00. Commuting peak can be considered as the major reason for peaks at 8:00 and 17:00. The reason for 0:00 will be explored in details later. Secondly, we assume that `Lighting Conditions`, `Road Descriptor`, and `Crash Descriptor` are highly correlated with the hour in a day. The following bar plots examine each variable independently with time. Also `Collision_Detail` will be used to see what the actual reasons for crashing in each hour. 

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Extract Columns for Time Analysis and Data Cleaning 
df_time_reason<- df %>% 
  mutate(time_hour = hour(Time)) %>%
  mutate(`Time` = factor(paste(as.character(time_hour), ":00", sep="")), 
         Lighting = `Lighting Conditions`,
         # ifelse(str_detect(`Lighting Conditions`, "Dark-Road"), 
         #                                sub("^\\S+\\s+","", `Lighting Conditions`), as.character(`Lighting Conditions`)), 
          RoadDescriptor = `Road Descriptor`, #abbreviate(`Road Descriptor`, 1, named = FALSE), 
         CrashDescriptor = `Crash Descriptor`, # abbreviate(`Crash Descriptor`, 1, named=FALSE), 
         CollisionTypeDescriptor = `Collision Type Descriptor`) %>%
  select(`Time`, time_hour, CrashDescriptor, Lighting, RoadDescriptor,  `Traffic Control Device`, Collision_Detail, Collision_Type, CollisionTypeDescriptor)

# dataframe for mosaic plots
t1 <-  df_time_reason %>%
  select(time_hour,Lighting) %>%
  mutate(Time = time_hour) %>%
  group_by(Time, Lighting) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  ungroup()

# t1$Lighting <- fct_relevel(t1$Lighting,"Unknown", after=Inf)
  
t2 <- df_time_reason %>%
  select(time_hour,RoadDescriptor) %>%
  mutate(Time = time_hour) %>%
  group_by(Time, RoadDescriptor) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  ungroup()
# t2$RoadDescriptor <- fct_recode(t2$RoadDescriptor, CG="CaG", CH="CaHC", CL="CaL", SH="SaHC", SG="SaG", SL="SaL", U="U")

t4 <- df_time_reason %>%
  select(time_hour,CrashDescriptor) %>%
  mutate(Time = time_hour) %>%
  group_by(Time, CrashDescriptor) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  ungroup()

t6 <- df_time_reason %>%
  select(time_hour,CollisionTypeDescriptor) %>%
  mutate(Time = time_hour) %>%
  group_by(Time, CollisionTypeDescriptor) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  ungroup()

```

**Lighting Condition**
<br>
The variation for lighting condition matches time. During day time, the majority cases are under daylight; and during evening, majority cases are under Dark-Road Lighted/Unlighted condition. For crashes happened in the night, the proportion of lighted condition are always higher than the proportion of unlighted condition. On the other side, 17:00 is when most crashes happened during one day, which is under daylight condition in most of the time. Therefore, we can simply conclude that the reason of evening crashes is not correlated with the lighting condition. 

```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}
fillcolors <- brewer.pal(6, "Set3")

t1$Lighting <- factor(t1$Lighting, levels = c("Unknown", "Daylight", "Dawn", "Dusk", "Dark-Road Lighted", "Dark-Road Unlighted"))

ggplot(t1, aes(x = factor(Time), y = RelFreq, fill = Lighting)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  ggtitle("Lighting Conditions of Car Crashes") + 
  labs(x = "Time", y = "Proportion") + 
  theme_classic()

# mosaic(Lighting~Time, direction=c("v", "h"), t1)
```

**Road Descriptor**
<br>
The assumption for road descriptor is that during night curve conditions may have more cases than straight conditions. However, the plot indicates that the road descriptor doesn't have a obvious correlation with time. In one day, most of the cases were under the straight and level condition, which is the major condition of most of the road in the states. 
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}
fillcolors <- brewer.pal(7, "Set3")

t2$RoadDescriptor <- fct_relevel(t2$RoadDescriptor,"Unknown")

ggplot(t2, aes(x = factor(Time), y = RelFreq, fill = RoadDescriptor)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  ggtitle("Road Descriptor of Car Crashes") + 
  labs(x = "Time", y = "Proportion") + 
  theme_classic()
# mosaic(RoadDescriptor~Time, direction=c("v", "h"), t2)
```

**Damage Severity **
<br>
Most of the crash in these two years are property damage accident. Injury accidentm, property damage & injury accident, and property damage accident have similar distribution through all the time during one day. The proportion of fatal accident is very small, but always happened during night and early morning. 
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}
fillcolors <- brewer.pal(4, "Set3")

t4$CrashDescriptor <- fct_relevel(t4$CrashDescriptor,"Fatal Accident", after=Inf)

ggplot(t4, aes(x = factor(Time), y = RelFreq, fill = CrashDescriptor)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  ggtitle("Damage Severity of Car Crashes") + 
  labs(x = "Time", y = "Proportion") + 
  theme_classic()
```


**Collision Details**

  There are total 30 levels for collision details, which is hard to illustrate in one plot. In the previous data cleaning part, collision details has a super class, collision type, which only have 3 levels ("Collision With", "Collision With Fixed Object", "Non-Collision"). The majority cases are under type "collision with", so the first plot only contains the collision details that belongs to type "collision with", and the second plot shows the top 6 levels in collision details for the remaining types "Non-Collision" and "Collision With Fixed Object". 


For collision with subject, motor vehicle collisions are over 50% of the cases, which doesn't have a obvious correlation with time. The second highest proportion is collision with animal, which has a peak in the morning around 5:00 and remains low proportion during day time. The animal activity and human activity can explain this phenomenon. Usually, animal activate when human activity reduces, which is during night and early morning. Other collision details, such as collision with bicyclist and pedestrain are not correlated with time. 
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}
h3 <- df_time_reason %>% 
  filter(Collision_Type == "Collision With" ) %>% #, Collision_Detail != "Other Motor Vehicle")%>%
  group_by(time_hour, Collision_Detail) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  ungroup()

h3$Collision_Detail <- h3$Collision_Detail %>%
  fct_relevel("Other Motor Vehicle", after=Inf)

fillcolors <- brewer.pal(7, "Set3") 

ggplot(h3, aes(x = factor(time_hour), y = RelFreq, fill = Collision_Detail)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  ggtitle("Collision Details of Car Crashes") + 
  labs(x = "Time", y = "Proportion") + 
  theme_classic()
```


As the plots shown below, details in both types are not have a correlation with the time. Details will be used in the follwing section to explore with other variables. 
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}
h2 <- df_time_reason %>% 
  filter(Collision_Type == "Non-Collision" ) %>%
  group_by(time_hour, Collision_Detail) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) 

h4 <- df_time_reason %>% 
  filter(Collision_Type == "Collision With Fixed Object") %>%
  group_by(time_hour, Collision_Detail) %>%
  summarise(Freq=n()) %>%
  top_n(6) %>%
  mutate(RelFreq = Freq / sum(Freq)) 

fillcolors <- brewer.pal(6, "Set3") 

g2 <- ggplot(h2, aes(x = factor(time_hour), y = RelFreq, fill = Collision_Detail)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  ggtitle("Non-Collision Details of Car Crashes") + 
  labs(x = "Time", y = "Proportion") + 
  theme_classic()

g4 <- ggplot(h4, aes(x = factor(time_hour), y = RelFreq, fill = Collision_Detail)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  ggtitle("Collision with Fixed Object Details of Car Crashes") + 
  labs(x = "Time", y = "Proportion") + 
  theme_classic()

layout <- "
A
B
"
wrap_plots(A = g2, B = g4, design = layout)
```


### Month of Year
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}
df %>% select(Date) %>%
  mutate(Month = month(Date)) %>%
  group_by(Month) %>%
  summarize(n = n()) %>%

ggplot(aes(factor(Month),n)) + 
  geom_col(fill="#35978F", alpha = 0.9) +
  labs(x="Month", y="Frequency") + 
  ggtitle("Car Crashes Occurred Each Month during 2018-2019") + 
  theme_minimal() 
```
    The bar plots above shows the number of car crashes occurred on each month accumulating 2018 and 2019. While the differences between each month are not tremendous, we could still observe that the number of crashes that happened in October, November, December and January are higher than the other months in a year. 
    Remarkably, the four months that have highest number of car crashes occurred are cluster together, and these four months covers almost the whole winter season. Considering the season factor, the weather conditions are different in winter than other season, and the daytime are shorter in winter which the lighting condition could be different as well. To get further information for car crashes that occurred in these four months, we will visulize through the mosiac plots between months and lighting conditions and weather conditions.

```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}

df_month <- df %>% 
  mutate(Month = month(Date))

df_light <- df_month %>%
  select(Month, `Lighting Conditions`) %>%
  group_by(Month, `Lighting Conditions`) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  ungroup()

fillcolors <- brewer.pal(6, "Blues")
df_light$`Lighting Conditions` <- factor(df_light$`Lighting Conditions`, levels = c("Unknown", "Daylight", "Dawn", "Dusk", "Dark-Road Lighted", "Dark-Road Unlighted"))
ggplot(df_light, aes(x = factor(Month), y = RelFreq, fill = `Lighting Conditions`)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  ggtitle("Lighting Conditions of Car Crashes") + 
  labs(x = "Month", y = "Proportion") + 
  theme_classic()

#mosaic(Lighting.Conditions ~ Month, direction=c("v", "h"), highlighting_fill = c("#F7FBFF", "#C6DBEF", "#6BAED6", "#4292C6", "#2171B5", "#084594"), df_light)

```


```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}
df_weather <- df_month %>%
  select(Month, `Weather Conditions`) %>%
  group_by(Month, `Weather Conditions`) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  ungroup()

fillcolors <- brewer.pal(8, "Blues")
df_weather$`Weather Conditions` <- factor(df_weather$`Weather Conditions`, levels = c("Unknown", "Other*", "Clear", "Cloudy", "Fog/Smog/Smoke", "Rain", "Sleet/Hail/Freezing Rain", "Snow"))
ggplot(df_weather, aes(x = factor(Month), y = RelFreq, fill = `Weather Conditions`)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  ggtitle("Weather Conditions of Car Crashes") + 
  labs(x = "Month", y = "Proportion") + 
  theme_classic()

```
    Looking at the two plots above, for the first plot, we can obverse that for the car crashes that occurred during October, November, December and January, the proportion of having lighting conditions of "Dark-Road Lighted" and "Dark-Road Unlighted" when crashes occurred are higher than those in the other months. While most of the crashes happened under the "Daylight" lighting condition, it is still worth noticing that how the proportion of lighting condition changed when the crashes occurred within these four months.
    In addition to seeing the lighting conditions, in the second plot showing the proportion of weather conditions when the car crashes occurred, we can observe that around 10 to 18 percent of car crashes occurred in November, December and January are under the "Snow" or "Sleet/Hail/Freezing Rain" weather condition. Since the weather condition vary objective to the time period (season) during a year, we cannot infer weather the higher number of crashes in these four months are due to change in the weather conditions.
    
## Where
```{r echo=FALSE, fig.height=20, fig.width=30, message=FALSE, warning=FALSE}
ny_county <- subset(map_data("county"), region == "new york") 
  
df_county <- right_join(df %>% 
                         group_by(`County Name`) %>%
                         summarise(n=n()) %>%
                         mutate(subregion = sub("\\.", "",tolower(`County Name`))) %>%
                         select(subregion, n), 
                       ny_county, 
                       by = "subregion") %>%
  select(subregion, n, long, lat, group) %>%
  drop_na()

lower_state <- df_county %>% 
  filter(n>50000) %>% 
  group_by(subregion) %>% 
  summarise(long=mean(long), lat=mean(lat))

embed_plot <- df_county %>% 
  filter(n>50000) %>%
ggplot(aes(x=long, y=lat)) +
  geom_polygon(aes(group = group, fill=n), color = "grey") +
  with(lower_state, annotate(geom="text", x = long, y = lat,
                             label=subregion, size=2)) +
  labs(title = "Long Island Details", x = "longitude", y = "latitude") +
  scale_fill_gradient(low="white", high="darkgreen") +
  theme(panel.background = element_blank()) 

ny_county_distinct <- ny_county %>% 
  group_by(subregion) %>% 
  summarise(long=mean(long), lat=mean(lat))

ggplot(data=df_county, aes(x=long, y=lat)) +
  geom_polygon(aes(group = group, fill=n), color = "grey") +
  with(ny_county_distinct, annotate(geom="text", x = long, y = lat, 
                                    label=subregion, size=3)) +
  labs(title = "New York State Crash by Counties", x = "longitude", y = "latitude") +
  scale_fill_gradient(low="white", high="darkgreen") +
  theme(panel.background = element_blank(), plot.title = element_text(hjust = 0.5)) +
  inset_element(embed_plot,0.01, 0.01, 0.35, 0.35)
```


## How

**Top 10 Crashes Detail**
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}

df_detail <- df %>% 
  group_by(Collision_Detail) %>%
  summarize(n = n()) %>%
  slice_max(order_by = n, n = 10)

ggplot(df_detail, aes(x = fct_reorder(Collision_Detail, n, .desc = TRUE), y=n)) + 
  geom_col(fill="Purple", alpha = 0.9) +
  labs(x="Crash Detail", y="Frequency") + 
  ggtitle("Top 10 Car Crashes Deatails") + 
  theme_minimal()

```
    Within the dataset, there is a variable providing the details of each crash case which give us the information of the type of collision. In total, there are 30 different situations. We pick the top 10 types which take over 80 percents of all crashes that occurred for analysis. According to the bar plot above, more than 70 percent of the crashes are collision with other motor vehicle and the following by collision with subject including Animal, Pedestrian, Bicyclist and fix objects such as Tree, Guide Rail and Sign Post.

**Road Surface patterns correspond to these top 10 type of Crashes**
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}

tenMost <- list(df_detail$Collision_Detail)

df_collision <- df_month %>%
  filter(Collision_Detail %in% df_detail$Collision_Detail) %>%
  select(Collision_Detail, `Road Surface Conditions`) %>%
  group_by(Collision_Detail, `Road Surface Conditions`) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  ungroup()

fillcolors <- brewer.pal(8, "Blues")
ggplot(df_collision, aes(x = RelFreq, y = fct_reorder(Collision_Detail, Freq, .desc = FALSE), fill = `Road Surface Conditions`)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  labs(x = "Proportion", y = "Collision Detail") +
  ggtitle("Road Surface Condition of the Corresponding Type of Crashes") + 
  theme_classic()

```

```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}

df_RoSur <- df %>% 
  group_by(`Road Surface Conditions`) %>%
  summarize(n = n())

ggplot(df_RoSur, aes(x = fct_reorder(`Road Surface Conditions`, n, .desc = TRUE), y=n)) + 
  geom_col(fill="DarkBlue", alpha = 0.9) +
  labs(x="Road Surface Conditions", y="Frequency") + 
  ggtitle("Road Surface Conditions When Car Crashes Occurred") + 
  theme_minimal()

```



**Combination: Lighting Conditions and Road Surface Condition**
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}

df_light <- df %>%
  select(`Road Surface Conditions`, `Lighting Conditions`) %>%
  group_by(`Road Surface Conditions`, `Lighting Conditions`) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  ungroup()

fillcolors <- brewer.pal(8, "Blues")
df_light$`Lighting Conditions` <- factor(df_light$`Lighting Conditions`, levels = c("Unknown", "Daylight", "Dawn", "Dusk", "Dark-Road Lighted", "Dark-Road Unlighted"))
ggplot(df_light, aes(x = `Road Surface Conditions`, y = RelFreq, fill = `Lighting Conditions`)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  ggtitle("Lighting and Road Surface Conditions of Car Crashes") + 
  labs(x = "Road Surface Condition", y = "Proportion") + 
  theme_classic()

```


**Traffic Control Condition**

```{r fig.width=8, echo=FALSE}


tenMost <- list(df_detail$Collision_Detail)

df_collision <- df_month %>%
  filter(Collision_Detail %in% df_detail$Collision_Detail, `Traffic Control Device` != "None") %>%
  select(Collision_Detail, `Traffic Control Device`) %>%
  group_by(Collision_Detail, `Traffic Control Device`) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  top_n(6) %>%
  ungroup()

fillcolors <- brewer.pal(12, "Set3")

ggplot(df_collision, aes(x = RelFreq, y = fct_reorder(Collision_Detail, Freq, .desc = FALSE), 
                         fill = `Traffic Control Device`)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  labs(x = "Proportion", y = "Collision Detail") +
  ggtitle("Road Surface Condition of the Corresponding Type of Crashes") + 
  theme_classic()
```