# Results

```{r message=FALSE, warning=FALSE, include=FALSE}
library(readr)
df <- read_csv("C:/Users/chaoy/Desktop/Fall2021/5702_EDAV/finalproject_data/cash_information_clean.csv")

#df <- read.csv("~/Downloads/cash_information_clean.csv")

#df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)], as.factor)
```

## When
```{r include=FALSE}
library(tidyverse)
library(dplyr)
library(lubridate)
library(vcd)
library(patchwork)
library(RColorBrewer)
```

### Time in Day
First, we explore when crashes happen the most, and the corresponding reasons. (Todo: x-axis label update to 0:00?)
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}
df_time <- df %>% select(Time, Year) %>%
  mutate(hour = hour(Time)) %>%
  group_by(hour, Year) %>%
  summarize(n = n()) %>%
  ungroup()

ggplot(df_time, aes(hour,n, fill=factor(Year))) + 
  geom_bar(stat="identity", position=position_dodge()) + 
  scale_fill_brewer(palette="Set2") +
  theme_minimal()
```
Here is an overview of number of crash occur in 24 hours in 2018 and 2019. Firstly, there are three peaks in the plot: 0:00, 8:00 and 17:00. Commuting peak can be considered as the major reason for peaks at 8:00 and 17:00. The reason for 0:00 will be explored in details later. Secondly, we assume that `Lighting Conditions`, `Road Descriptor`, `Traffic Control Device`, `Crash Descriptor` and `Collision Type Descriptor` are highly correlated with the hour in a day. The following mosaic plots and histograms will examine each variable independent with time. Also `Collision_Detail` will be used to see what the actual reasons for crashing in each hour. 

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Extract Columns for Time Analysis and Data Cleaning 
df_time_reason<- df %>% 
  mutate(time_hour = hour(Time)) %>%
  mutate(`Time` = factor(paste(as.character(time_hour), ":00", sep="")), 
         Lighting = ifelse(str_detect(`Lighting Conditions`, "Dark-Road"), 
                                        sub("^\\S+\\s+","", `Lighting Conditions`), as.character(`Lighting Conditions`)), 
          RoadDescriptor = abbreviate(`Road Descriptor`, 1, named = FALSE), 
         CrashDescriptor = abbreviate(`Crash Descriptor`, 1, named=FALSE), 
         CollisionTypeDescriptor = `Collision Type Descriptor`) %>%
  select(`Time`, time_hour, CrashDescriptor, Lighting, RoadDescriptor,  `Traffic Control Device`, Collision_Detail, Collision_Type, CollisionTypeDescriptor)

# dataframe for mosaic plots
t1 <-  df_time_reason %>%
  select(time_hour,Lighting) %>%
  mutate(Time = time_hour) %>%
  group_by(Time, Lighting) %>%
  summarise(Freq=n()) %>%
  ungroup()
t1$Lighting <- fct_relevel(t1$Lighting,"Unknown", after=Inf)
  
t2 <- df_time_reason %>%
  select(time_hour,RoadDescriptor) %>%
  mutate(Time = time_hour) %>%
  group_by(Time, RoadDescriptor) %>%
  summarise(Freq=n()) %>%
  ungroup()
t2$RoadDescriptor <- fct_recode(t2$RoadDescriptor, CG="CaG", CH="CaHC", CL="CaL", SH="SaHC", SG="SaG", SL="SaL", U="U")

t4 <- df_time_reason %>%
  select(time_hour,CrashDescriptor) %>%
  mutate(Time = time_hour) %>%
  group_by(Time, CrashDescriptor) %>%
  summarise(Freq=n()) %>%
  ungroup()

t6 <- df_time_reason %>%
  select(time_hour,CollisionTypeDescriptor) %>%
  mutate(Time = time_hour) %>%
  group_by(Time, CollisionTypeDescriptor) %>%
  summarise(Freq=n()) %>%
  ungroup()
t6$CollisionTypeDescriptor <- fct_recode(t6$CollisionTypeDescriptor, H="HEAD ON", `0`="LEFT TURN (0)", `3`="LEFT TURN (3)", 
                                         O="OTHER", OT="OVERTAKING", RE="REAR END", RA="RIGHT ANGLE", `5`="RIGHT TURN (5)", 
                                         `6`="RIGHT TURN (6)", S="SIDESWIPE", U="Unknown")
```

**Lighting Condition**
```{r fig.width=15, fig.height=10, message=FALSE, warning=FALSE, echo=FALSE}
mosaic(Lighting~Time, direction=c("v", "h"), t1)
```

**Road Surface Condition**
```{r fig.width=15, fig.height=10, message=FALSE, warning=FALSE, echo=FALSE}
mosaic(RoadDescriptor~Time, direction=c("v", "h"), t2)
```

**Demage Severity **
```{r fig.width=15,fig.height=10, message=FALSE, warning=FALSE, echo=FALSE}
mosaic(CrashDescriptor~Time, direction=c("v", "h"), t4)
```
**Collision Type Description**

```{r fig.width=15, fig.height=10, message=FALSE, warning=FALSE, echo=FALSE}
mosaic(CollisionTypeDescriptor~Time, direction=c("v", "h"), t6)
```


**Traffic Control Condition**
```{r fig.width=10, fig.height=10, echo=FALSE}
h1 <- df_time_reason %>% 
  group_by(time_hour, `Traffic Control Device`) %>% summarise(n=n()) %>%
  top_n(6) %>% ungroup()

h1$`Traffic Control Device` <- h1$`Traffic Control Device` %>%
  fct_relevel("None") %>% 
  fct_recode(NP="No Passing Zone", None="None", Other="Other", Stop="Stop Sign", Signal="Traffic Signal", Y="Yield Sign", U="Unknown")

ggplot(h1, aes(x=`Traffic Control Device`, y=n)) + geom_col() + facet_wrap(~time_hour, nrow=6, ncol=4)
```


**Collision Details**
```{r fig.width=10, fig.height=10, message=FALSE, warning=FALSE, echo=FALSE}
h2 <- df_time_reason %>% 
  filter(Collision_Type != "Collision With" ) %>%
  group_by(time_hour, Collision_Detail) %>%
  summarise(n=n()) %>%
  top_n(5) %>%
  ungroup()

h2$Collision_Detail <- h2$Collision_Detail %>% fct_recode(E="Earth Embankment/Rock Cut/Ditch", G="Guide Rail - Not At End", Tree="Tree",
                                                          U="Light Support/Utility Pole", Other="Other Fixed Object*", Sign="Sign Post") %>%
  fct_relevel("Other", "Sign", after=Inf)

ggplot(h2, aes(x=factor(Collision_Detail), y=n)) + 
geom_col(fill="#8da675")+ 
facet_wrap(~time_hour, nrow=6, ncol=4)+ 
labs(x="", y="", title="Collision Details vs Time")
```


```{r fig.width=10, fig.height=10, message=FALSE, warning=FALSE, echo=FALSE}
h3 <- df_time_reason %>% 
  filter(Collision_Type == "Collision With" ) %>% #, Collision_Detail != "Other Motor Vehicle")%>%
  group_by(time_hour, Collision_Detail) %>%
  summarise(n=n()) %>%
  ungroup()

h3$Collision_Detail <- h3$Collision_Detail %>% fct_recode(Ani="Animal", Bike="Bicyclist", S="In-Line Skater", Vehicle="Other Motor Vehicle",
                                  O="Other Object (Not Fixed)*", P="Pedestrian", Train="Railroad Train") %>%
  fct_relevel('Vehicle', 'Ani','P')

ggplot(h3, aes(x=factor(Collision_Detail), y=n)) + 
geom_col(fill="#8da675")+ 
facet_wrap(~time_hour, nrow=6, ncol=4)+ 
labs(x="", y="", title="Collision Details vs Time")
```


### Day of Week
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}
df %>%
  select(`Day of Week`, Year) %>%
  group_by(`Day of Week`, Year) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  
ggplot(aes(fct_relevel(`Day of Week`, "Monday", "Tuesday", "Wednesday","Thursday", "Friday", "Saturday", "Sunday"),n,  fill=factor(Year))) +
  geom_bar(stat="identity", position=position_dodge()) + 
  scale_fill_brewer(palette="Set2") +
  labs(x="Day of Week", y="Frequency") +
  theme_minimal()
```

### Month of Year
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}
df %>% select(Date) %>%
  mutate(Month = month(Date)) %>%
  group_by(Month) %>%
  summarize(n = n()) %>%

ggplot(aes(factor(Month),n)) + 
  geom_col(fill="#35978F", alpha = 0.9) +
  labs(x="Month", y="Frequency") + 
  ggtitle("Car Crashes Occurred Each Month during 2018-2019") + 
  theme_minimal() 
```
    The bar plots above shows the number of car crashes occurred on each month accumulating 2018 and 2019. While the differences between each month are not tremendous, we could still observe that the number of crashes that happened in October, November, December and January are higher than the other months in a year. 
    Remarkably, the four months that have highest number of car crashes occurred are cluster together, and these four months covers almost the whole winter season. Considering the season factor, the weather conditions are different in winter than other season, and the daytime are shorter in winter which the lighting condition could be different as well. 

```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}

df_month <- df %>% 
  mutate(Month = month(Date))

df_light <- df_month %>%
  select(Month, Lighting.Conditions) %>%
  group_by(Month, Lighting.Conditions) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  ungroup()

fillcolors <- brewer.pal(6, "Blues")
df_light$Lighting.Conditions <- factor(df_light$Lighting.Conditions, levels = c("Unknown", "Daylight", "Dawn", "Dusk", "Dark-Road Lighted", "Dark-Road Unlighted"))
ggplot(df_light, aes(x = factor(Month), y = RelFreq, fill = Lighting.Conditions)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  ggtitle("Lighting Conditions of Car Crashes") + 
  labs(x = "Month", y = "Proportion") + 
  theme_classic()

#mosaic(Lighting.Conditions ~ Month, direction=c("v", "h"), highlighting_fill = c("#F7FBFF", "#C6DBEF", "#6BAED6", "#4292C6", "#2171B5", "#084594"), df_light)

```


## Where
```{r echo=FALSE, fig.height=20, fig.width=30, message=FALSE, warning=FALSE}
ny_county <- subset(map_data("county"), region == "new york") 
  
df_county <- right_join(df %>% 
                         group_by(`County Name`) %>%
                         summarise(n=n()) %>%
                         mutate(subregion = sub("\\.", "",tolower(`County Name`))) %>%
                         select(subregion, n), 
                       ny_county, 
                       by = "subregion") %>%
  select(subregion, n, long, lat, group) %>%
  drop_na()

lower_state <- df_county %>% 
  filter(n>50000) %>% 
  group_by(subregion) %>% 
  summarise(long=mean(long), lat=mean(lat))

embed_plot <- df_county %>% 
  filter(n>50000) %>%
ggplot(aes(x=long, y=lat)) +
  geom_polygon(aes(group = group, fill=n), color = "grey") +
  with(lower_state, annotate(geom="text", x = long, y = lat,
                             label=subregion, size=2)) +
  labs(title = "Long Island Details", x = "longitude", y = "latitude") +
  scale_fill_gradient(low="white", high="darkgreen") +
  theme(panel.background = element_blank()) 

ny_county_distinct <- ny_county %>% 
  group_by(subregion) %>% 
  summarise(long=mean(long), lat=mean(lat))

ggplot(data=df_county, aes(x=long, y=lat)) +
  geom_polygon(aes(group = group, fill=n), color = "grey") +
  with(ny_county_distinct, annotate(geom="text", x = long, y = lat, 
                                    label=subregion, size=3)) +
  labs(title = "New York State Crash by Counties", x = "longitude", y = "latitude") +
  scale_fill_gradient(low="white", high="darkgreen") +
  theme(panel.background = element_blank(), plot.title = element_text(hjust = 0.5)) +
  inset_element(embed_plot,0.01, 0.01, 0.35, 0.35)
```


## How

**Top 10 Crahes Detail**
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}

df_detail <- df %>% 
  group_by(Collision_Detail) %>%
  summarize(n = n()) %>%
  slice_max(order_by = n, n = 10)

ggplot(df_detail, aes(x = fct_reorder(Collision_Detail, n, .desc = TRUE), y=n)) + 
  geom_col(fill="Purple", alpha = 0.9) +
  labs(x="Crash Detail", y="Frequency") + 
  ggtitle("Top 10 Car Crashes Deatails") + 
  theme_minimal()

```


**Road Surface patterns correspond to these top 10 type of Crashes**
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}

tenMost <- list(df_detail$Collision_Detail)

df_collision <- df_month %>%
  filter(Collision_Detail %in% df_detail$Collision_Detail) %>%
  select(Collision_Detail, Road.Surface.Conditions) %>%
  group_by(Collision_Detail, Road.Surface.Conditions) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  ungroup()

fillcolors <- brewer.pal(8, "Blues")
ggplot(df_collision, aes(x = RelFreq, y = fct_reorder(Collision_Detail, Freq, .desc = FALSE), fill = Road.Surface.Conditions)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  labs(x = "Proportion", y = "Collision Detail") +
  ggtitle("Road Surface Condition of the Corresponding Type of Crashes") + 
  theme_classic()

```

```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}

df_RoSur <- df %>% 
  group_by(Road.Surface.Conditions) %>%
  summarize(n = n())

ggplot(df_RoSur, aes(x = fct_reorder(Road.Surface.Conditions, n, .desc = TRUE), y=n)) + 
  geom_col(fill="DarkBlue", alpha = 0.9) +
  labs(x="Road Surface Conditions", y="Frequency") + 
  ggtitle("Road Surface Conditions When Car Crashes Occurred") + 
  theme_minimal()

```



**Combination: Lighting Conditions and Road Surface Condition**
```{r fig.width=8, message=FALSE, warning=FALSE, echo=FALSE}

df_light <- df %>%
  select(Road.Surface.Conditions, Lighting.Conditions) %>%
  group_by(Road.Surface.Conditions, Lighting.Conditions) %>%
  summarise(Freq=n()) %>%
  mutate(RelFreq = Freq / sum(Freq)) %>%
  ungroup()

fillcolors <- brewer.pal(8, "Blues")
df_light$Lighting.Conditions <- factor(df_light$Lighting.Conditions, levels = c("Unknown", "Daylight", "Dawn", "Dusk", "Dark-Road Lighted", "Dark-Road Unlighted"))
ggplot(df_light, aes(x = Road.Surface.Conditions, y = RelFreq, fill = Lighting.Conditions)) + 
  geom_col() + 
  scale_fill_manual(values = fillcolors) + 
  ggtitle("Lighting and Road Surface Conditions of Car Crashes") + 
  labs(x = "Road Surface Condition", y = "Proportion") + 
  theme_classic()

```

